---
title: "Module 04"
author: "Victoria Zdanowicz"
date: "9/10/2021"
output: 
  html_document:
    toc: TRUE
    toc_depth: 3
    toc_float: TRUE
---

Below is my work for Module 4 
I first created a text file (with Atom) of the random-people.csv data and loaded the 'Stringr' package.
<br>
```{r}
library(stringr)
library(knitr)
```
<br><br>

# CHALLENGE 1

<br><br>
First, I removed the punctuation/special characters from the opening line to James Joyce's Ulysses, “Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed.” I then used str_split() to create a vector of 1, word character strings.
<br>
```{r}
string<-"Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed."
gsub("[[:punct:]]","",string)
z<-gsub("[[:punct:]]","",string)
str_split(z, boundary("word"))
```
<br>
The next instruction was to extract every third element in the vector. I first created a matrix with the character strings from the previous step. I then extracted every third word from the vector by commanding a sequence with every third element.
<br>
```{r}
x<-matrix(data= c("Stately","plump","Buck","Mulligan","came","from","the","stairhead","bearing","a","bowl","of","lather","on","which","a","mirror","and","a","razor","lay","crossed"),nrow = 2,ncol = 11, byrow=FALSE)
y <- x[seq(1, length(x), 3)]
y
```
<br><br>

***

<br><br>

# CHALLENGE 2

<br><br>
I first created a matrix with values 1-80. 
<br>
```{r}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
m
```
<br>
Given the matrix, m, above, I extracted the 2nd, 3rd, and 6th columns and assigned them to the variable x
<br>
```{r}
x<-m[,c(2,3,6)]
x
```
<br>
Given the same matrix, m, I extracted the 6th to 8th row and assign them to the variable x
<br>
```{r}
x<-m[c(6:8),]
x
```
<br>
Given the same matrix, m, I extracted the elements from row 2, column 2 to row 6, column 9 and assign them to the variable x
<br>
```{R}
x<-m[c(2:6),c(2:9)]
x
```
<br><br>

***

<br><br>

# CHALLENGE 3

<br><br>
I first constructed a 4-dimensional, 400 element array (5 x 5 x 4 x 4) named a, consisting of the numbers 400 to 1 
<br>
```{r}
a <- array(400:1, c(5, 5, 4, 4))
a
```
<br>
With this matrix, a[1, 1, 1, 2] returns the element located at [1,1] in the fifth 5X5 matrix [, , 1, 2]
<br>
```{r}
a[1, 1, 1, 2]
```
<br>
With this matrix, a[2, 3, 2, ] returns the elements located at [2,3] in [, , 2, ] matrices (matrices with 2 in the second to last position?) I am not sure how to describe or refer to a specific matrix/vector(?) within this array. Matrix/array 1-16?
<br>
```{r}
a[2, 3, 2, ]
```
<br>
With the same matrix, a[1:5, 1:5, 3, 3] returns the elements from [1,1] to [5,5] in matrix 11 [, , 3, 3], which is the entire matrix (again, I am unsure how to refer to that specific matrix/array/vector?)
<br>
```{r}
a[1:5, 1:5, 3, 3]
```
<br><br>

***

<br><br>

# Overwriting

<br><br>
Given matrix, m, from earlier, below are my explanations/descriptions of the various provided operations
<br>
m[7,1] <- 564
	This changes the element at [7,1] to new value 564
```{r}
m[7, 1] <- 564
print(m)
```
<br><br>
m[, 8] <- 2
	This changes all elements in column 8 to new value, 2
```{r}
m[, 8] <- 2
print(m)
```
<br><br>
m[2:5, 4:8] <- 1
	This changes all elements in rows 2:5 that are within columns 4:8 to value, 1
```{r}
m[2:5, 4:8] <- 1
print(m)
```
<br><br>
m[2:5, 4:8] <- c(20, 19, 18, 17)
	This changes the elements in [2,4:8] to 20; the elements in [3,4:8] to 19; the elements in[4,4:8] to 18, and the elements in [5,4:8] to 17 
```{r}
m[2:5, 4:8] <- c(20, 19, 18, 17)
print(m)
```
<br><br>
m[2:5, 4:8] <- matrix(data = c(20:1), nrow = 4, ncol = 5, byrow = TRUE)
	This changes the elements in rows 2:5 that are within columns 4:8 to descending sequence from 20 to 1, with values changes by row instead of default by column,creating a 4X5 matrix with descending values 20:1
```{r}
m[2:5, 4:8] <- matrix(data = c(20:1), nrow = 4, ncol = 5, byrow = TRUE)
print(m)
```
<br><br>
m[,8] <- c(“a”, “b”)
	This alters the elements in row 8 [,8] to alternating characters “a” and “b”
```{r}
m[, 8] <- c("a", "b")
print(m)
```
<br><br>

***

<br><br>

# CHALLENGE 4

<br><br>
To create a list of simplified primate taxonomy, I began creating lists going from most specific taxonomic rank to most general, starting with the suborder Strepsirhini 
<br>
```{r}
Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae",
    "Daubentoniidae")
Lorisoidea <- c("Lorisidae", "Galagidae")
Strepsirhini <- list(Lemuroidea, Lorisoidea)
names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea")
Strepsirhini
```
<br><br>
I then followed this same process to create 2 lists within the suborder Haplorhini starting with the parvorders Platyrrhini and Catarrhini within the infraorder Anthropoidea
<br>
```{r}
Platyrrhini <- c("Cebidae", "Atelidae", "Pitheciidae")
Catarrhini <- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Anthropoidea <- list(Platyrrhini, Catarrhini)
names(Anthropoidea) <- list("Platyrrhini", "Catarrhini")
Anthropoidea
```
<br><br>
Next I created a list of the infraorder Tarsioidea to combine with the infraorder Anthropoidea list for the suborder Haplorhini
<br>
```{r}
Tarsioidea <- c("Tarsiidae")
Haplorhini <- list(Tarsioidea, Anthropoidea)
names(Haplorhini) <- list("Tarsioidea", "Anthropoidea")
Haplorhini
```
<br><br>
Lastly, I combined the Haplorhini and Strepsirhini lists to create the overall Primate list 
<br>
```{r}
Primates <- list(Haplorhini, Strepsirhini)
names(Primates) <- list("Haplorhini", "Strepsirhini")
Primates
```
<br><br>

***

<br><br>

# CHALLENGE 5

<br><br>
I first created a 5 X 3 matrix, m3, with the values: 3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2
<br>
```{R}
m3<-matrix(data = c(3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2), nrow = 5, ncol = 3, byrow = TRUE)
m3
```
<br><br>
Next - I coerced the matrix into a data frame, m3.df, with the command as.data.frame()
<br>
```{R}
m3.df <- as.data.frame(m3)
m3.df
```
<br><br>
To coerce the second column to be logical-valued I used the as.logical() command 
<br>
```{R}
m3.df$V2<-as.logical(m3.df$V2)
m3.df
```
<br><br>
To coerce the third column to be factor-valued I used the factor() command 
<br>
```{R}
m3.df$V3<-as.factor(m3.df$V3)
m3.df
```
<br><br>
Lastly, the str() command displays the data type for each variable in m3.df
<br>
```{R}
str(m3.df)
```
<br><br><br><br><br><br>