---
title: "Module 4 Challenges"
author: "Victoria Zdanowicz"
date: "9/10/2021"
output: html_document
---

Below is my work for the Module 4 Challenges. 
I first created a text file (with Atom) of the random_people.csv data and loaded the 'Stringr' package.
```{r}
library(stringr)
library(knitr)
```

**CHALLENGE 1**

First, I removed the punctuation/special characters from the opening line to James Joyce's *Ulysses*, “Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed.” I then used `r str_split()` to create a vector of 1, word character strings.

```{r}
string<-"Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed."
gsub("[[:punct:]]","",string)
z<-gsub("[[:punct:]]","",string)
str_split(z, boundary("word"))
```

The next instruction was to extract every third element in the vector. I first created a matrix with the character strings from the previous step. I then extracted every third word from the vector.

```{r}
m<-matrix(data= c("Stately","plump","Buck","Mulligan","came","from","the","stairhead","bearing","a","bowl","of","lather","on","which","a","mirror","and","a","razor","lay","crossed"),nrow = 2,ncol = 11, byrow=FALSE)
y <- m[seq(1, length(m), 3)]
y
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}

```




